---
title: "tvarGIRF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tvarGIRF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`tvarGIRF` is an `R` package that calculates generalised impulse response functions to reduced form shocks for threshold vector autoregressions estimated using the `tsDyn` package.

**Please note**: this package is something I put together quickly for an uncompleted project because the only other implementation of GIRFs in R [that existed at the time](http://groups.google.com/group/tsdyn/t/5c517a94a3a3ab0c) looks like it implements the bootstrap incorrectly to me. According to the `tsDyn` [wiki](https://github.com/MatthieuStigler/tsDyn/wiki/FAQ#1-are-generalized-impulse-functions-girf-available-in-tsdyn), GIRFs are now included in the `tsDyn` package, so you might have better luck with those.

## Installation

Install the lastest released version of the package using the R `remotes` package:
```{r eval = FALSE}
library(remotes)
install_github("angusmoore/tvarGIRF", ref = "v0.1.2")
```

You can also install the latest development version:
```{r eval = FALSE}
library(remotes)
install_github("angusmoore/tvarGIRF")
```

## Basic usage

The library extends the R `tsDyn` package. The following example illustrates how
to create a simple GIRF for a threshold VAR using the `zeroyld` dataset provided
with the `tsDyn` package.

`GIRF` is given a reduced form shock - in the example below a shock to only the
second variable `c(0,1)`. It is up to you to identify shocks;
`tvarGIRF` won't do it for you.

```{r warning=FALSE}
library(tsDyn)
library(tvarGIRF)

# Estimate an example TVAR using the zeroyld dataset included in the tsDyn package
data(zeroyld)
exampleTVAR <- TVAR(zeroyld, lag=2, nthresh=1, thDelay=1, mTh=1, plot=FALSE)

# Calculate GIRFs for a reduced form shock to the second variable (long.run)
GIRF(exampleTVAR, c(0,1), H = 10, R = 10)
```

NB: These GIRFs are meaningless as I have used very few repetitions to 
calculate them. I have done this so that the examples don't take too long to run.
But you should use _much_ larger values for your actual estimation.

### Shocks to more than one variable

Note that you can shock more than one variable. If you had a three equation TVAR,
passing `shock = c(1,0,0)` would give you the GIRF to a reduced-form shock to the
first equation, and `shock = c(1,1,0)` would be a reduced form shock to both the
first and second. The values need not be 1 either. You can estimate the response
to a 0.9 unit shock to one variable and a 0.645 to the other (just as an example!):
```{r}
GIRF(exampleTVAR, c(0.9, 0.645), H = 10, R = 10)
```
This is helpful, as you will typically want to identify _structural_ shocks. The
reduced form shock will be some linear combination of those structural shocks. See
the section "A note on shocks" for a bit more discussion.

### Viewing the results

`tvarGIRF` has a couple helpful functions to help you view the results from `GIRF`.

The first is to simply print the results of an estimation:
```{r}
saved_girfs <- GIRF(exampleTVAR, c(0,1), H = 10, R = 10)
saved_girfs # or, equivalently print(saved_girfs)
```

There is also a tidier for GIRFs (a la [`broom`](https://cran.r-project.org/web/packages/broom/vignettes/broom.html)):
```{r}
tidy(saved_girfs)
```

You can `View` GIRFs in the spreadsheet viewer. You can also plot your GIRFs, using
a built in `plot` function that uses `ggplot2`:
```{r}
plot(saved_girfs)
```

(Again, these GIRFs are meaningless as I have used very few repetitions to 
calculate them, which is why they look so wonky. I have done this so that the
examples don't take too long to run. But you should use _much_ larger values for
your actual estimation.)

## A note on shocks

`tvarGIRF` **does not identify shocks for you**. The argument you supply to tvarGIRF
is for a reduced form shock. This is rarely what you want. The reduced form shocks
are extremely unlikely to be uncorrelated with one another, and so it makes little
sense to consider one such isolated shock.

Instead, you will want to identify _structural_ shocks that _are_ uncorrelated
with one another.

If you want to use structural shocks, you will have to identify the shocks yourself
and supply the reduced form shock that corresponds to your chosen structural shock.
**It is up to you to identify shocks; `tvarGIRF` won't do it for you.** There are
many strategies for identifying structural shocks, `tvarGIRf` can't help with this.

As an example, let's identify a structural shock for the TVAR we estimated above
using timing restrictions / Cholesky decomposition. Our timing assumption is that
shocks to `short.run` affect both variables contemparenously, but shocks to
`long.run` affect `short.run` with a one period lag.^[This is not a very sensible
assumption in this context and is meant purely to illustrate how you might go
about this identification strategy using `tvarGIRF`. There are many problems
with the identification strategy I use here, including that I assume that the
impact of the structural shocks is the same in both regimes.]

Using our estimated TVAR, we can get the residuals by:
```{r}
e <- resid(exampleTVAR)
```

We calculate the covariance matrix by:
```{r}
cov_matrix <- t(e) %*% e
print(cov_matrix)
```
And take an upper triangular Cholesky decomposition:
```{r}
chol(cov_matrix)
```
I could then calculate the GIRF to the identified structural shock to the variable
`short.run` by running:
```{r}
structural_girfs <- GIRF(exampleTVAR, c(6.535195, 8.393587), H = 10, R = 10)
```
Which is what my identification scheme tells me the structural shock looks like
as a reduced form shock (the first row of the cholesky-decomposed covariance matrix).

## GIRFs conditional on starting in a particular regime

The `restrict.to` argument takes an integer corresponding to which regime you wish
to restrict the GIRF to. For instance, using the toy example we started with before,
if I estimate the following TVAR:
```{r}
data(zeroyld)
exampleTVAR <- TVAR(zeroyld, lag = 2, nthresh = 1, thDelay = 1, mTh = 1, plot = FALSE)
```
I have a TVAR with _two_ regimes. These are regimes are denoted by `1` when then
threshold variable (in this case `short.run`) is below my estimated threshold
(`10.653` in this example) and denoted by `2` when it is above the threshold value.
This scales to more regimes: `1` corresponds to the lowest regime and `nthresh+1`
to the highest regime.

You can see the names of your regimes by `print`ing your TVAR:
```{r}
exampleTVAR
```
My regimes are called `Bdown` (the regime denoted by 1, as it appears first) and
`Bup` (the regime denoted by 2, as it appears second).

You can also see which regime your estimated model is in at each point in your
estimation sample by looking at:
```{r}
regime(exampleTVAR)
```
And you could confirm that these correspond to `short.run` being above or below
the estimated threshold value.

The `restrict.to` argument in tvarGIRF uses those regime numbers. Suppose I want to
examine GIRFs conditional on starting in regime `Bdown`, when `short.run` is below
my threshold. This corresponds to regime `1`. So I would use `restrict.to = 1` in
my call to GIRF:
```{r}
GIRF(exampleTVAR, c(0,1), H = 10, R = 10, restrict.to = 1)
```

(NB: Again, I am using far too few repetitions to calculate meaningful GIRFs, but
I don't want the examples to take too long.)
